### 2020-06-18 11-25/todo.tex

\documentclass{article}

\begin{document}
\title{Week Nov 19-23}

\maketitle
\section{To Do?}

%idk haha \\
%outline natin mga things to do from the minutes \\
%try to fix what we can pero no pressure \\
Intro \\ \\
- Lessen Bio Slides \\
- Explain client server \\ 
- Have more tech slides \\
- Explain Database core principles CIA? \\
- APIs? \\
- Add to P2P explain it more \\
- Find out all jargon used \\
- After that define those jargon \\
- Describe Distributed Databases more \\
- add more pictures \\
- explain Client Server and Peer 2 Peer \\
- explain network layers \\

Problem \\ \\
- Describe the problem why genomic researchers make these systems in the first place \\
- Describe problem where researchers download the whole database kasi mabagal internet nila or something \\

RRL \\ \\
- may one table comparing all the systems may check and ex \\

Objectives \\ \\
- Add some way of measuring objectives \\
- More concise \\
- Example: Implement an info delivery system that speedups download of data \\

\section{NOTES}
\subsection{Problem Statement}

\subsection{Theoretical Framework}
from seqtorr
POST
PUT
GET
DELETE

POST
1) choose file
2) turn file to torrent file
3) upload torrent file

PUT
1) choose metadata to edit
2) edit metadata
3) save metadata

GET (may expound the system more)
1) get torrent file
2) from torrent file get IP address of peers and parts they have
3) measure speed of each part from each peer
4) download the part from the fastest peer
5) combine files into the file
6) checksum??

DELETE
1) delete torrent?

\subsection{Proposed Methodology}
CREATING THE SYSTEM \\
1) code pseudocode / prototype for UI \\
2) code network shit / test out network shit \\
3) code base code \\

4) code frontend \\
5) deploy \\

TESTING \\
6) create a test for testing original speeds \\
7) test original speeds \\
8) create a test for testing new speeds \\
9) test new speeds \\
10) test UI??? \\
11) test security?? \\
12) test checksum???? \\

%Ideas kaya comment pa
%Theoretical Framework
% main server
%

%Proposed Methodology
% Mid Jan - Mid April
% Per Week?

\subsection{Saturday}
10pm - 12 mn \\
revise presentation \\
12mn - 2am \\
start theoretical framework \\
start proposed methodology \\

\subsection{Sunday}

9am-6pm wala si cid pero Carl you can start

\subsection{Sunday}
let's make a goal kaya na rest day natin sunday (and monday morning) para di tayo

\end{document}