### 2020-06-18 12-09/minutes.tex

scalability
user interface
upload speed

specific objectives:
benchmarking objectives address those objectives

??accuracy, ??error rate

download conditions?

can test if a data node goes down? robustness

raw download speeds? is there a standard way of downloading that incorporates the network used.

properly define metrics na

ano yung expected type of environment we would be testing our system? 
expected: test it on natural environment
use network on ny given day?

controlled network infrastructure? or to show unperturbed download speed

NCBI, there is another downloading way of whole database download

and use that for benchmarking

how much resources do you see.

resources:
- set up a workstation, gawing master server, won't be disturbing for others
- specify more the projected resources we need
: clear is one master na always
: one data node sa baba
: one sa dcs
: ask sir cabarle, sir mario, a private network
: public, tapos firewall IP table block,
: Cpanel ...

revised proposal:
- include resources needed
- put it in methodology

comm. assisted. dist. database for sequences

caDDS

- PGC data node, is a workstation sa baba, within the network.
- if testing outside, DCS.

WED Dec 11

FRIDAY Dec 13, 12 nn. Official Rev. Proposal na Paper

FINAL FINAL DEADLINE BY 18 12 nn dapat satisfactory na.

Kung may revision, incomplete pa yun.