### 2020-06-18 Paper2020ACM/meeting 1-21-2020.tex

199-1-21 Meet Notes
Data sync bet servers

when a dataset gets uploaded:
- notify master node of new update
- master picks which other data nodes will host a copy of the dataset
- then asks those nodes to copy from the original data node
- data nodes will copy from the original
- after complete download, flag locally and notify master node about completion

upd or delete:
- inform everyone if metadata has been changed

only master node keeps a record of all metadata in the dist system
all data nodes have to ask master node if user wants a file it does not own

use nginx
http://w.x.y.z:80/api/ (exposed endpoint) -> http://w.x.y.z:5984 (couchdb)

assume data nodes have a public ip
assume all data is public

dev:
start coding NOW, then decide what features to change/add/remove

benchmarking:
refer to lit on factors to consider for benchmarking
look for fasta files