### 2020-06-26 Paper2020ACM/MeetingJune25.tex
June 25 9pm meeting notes

Marami to hahahaha get ready mostly sa proof of correctness

Whole Paper:
1) Make it 2 column format
2) Add sir El King Morado - UP Philippine Genome Center
3) Remove references to page numbers
4) Put figures on top of a column (see what sir el king sent)

Abstract
- okay content
- fix typoL Data nods

Intro Pg 1
- Remove sentence: [the goal of this intro]
- remove subsections in inntro
- Genomics study of blank. To generate data you need sequencing. But the data is big & needs effort. To sequence a lot, tech evolved and NGS gave high throughput
- Intro should be 2 columns lang. Make each paragraph a few sentences instead.
- Intro goal: layoutht background of the work
- Flow: Idea of A lot of data -> Problems -> Proposed Solution
-

-Intro Pg 2
- 3,000,000,000 -> 3 billion
- Shorten HGP section. Combine HGP + NGS to one paragraph.
- Fix typo: Genomics is is

-Intro: Database
- NGS -> high voluminous data -> need to store it in a digital format -> retrieve it for analysis

Overall info guide: 
- tama na yung flow
- lessen info
- compress database to 1/2 sentences. Then mention centralized/distributed database.

Into: CAP
- make it 1 paragraph
- only describe the most important principle and why

Application Architecture: (Move to Theoretical Framework)
- Discuss Client Server in 1 sentence
- Discuss P2P in 1 sentence
-Describe Hybrid

-Sample: 
- First we need to discuss 2 basic architecture: CS, P2P. 
- CS is blank. 
- P2P is blank. 
- Our proposed system will be hybrid getting these aspects from CS and P2P

RRL:
- first paragraph say: We review blank. We would like to focus on blank and blank. In the end we would like to make a comparitive analysis. 
- Table Comparing the diff Databass: Remove Dat, Describe each column narratively, and describe some conclusions from the table
- Answer the question: anong halaga ng columns? 
- Discuss each column CS/P2P/H, and the data on each column
- FASTA -> Biological
- Move dat to Methodology.

Review of Related Literature:
- the review tells you you are meant to critique the work you have seen
- here summarize gaps (mej copy the problems to here)

Problem:
- summarize the gaps. To review the current implementation has problems in size.
- To summarize, centralized databases suffer from these problems. Therefore the research aims to focus on developing a database with these features
- Problem Statement: Central thesis, one unifying question. May certain level of measurability.

Objectives:
- specify metrics, activities, na once you do this succeed.
- the bullet points "Resolve problems blank of blank" move them to a paragraph below objectives
- For the paragraph below: First objective addresses these problems, second objective addresses these limitations blank.
- keep first 2 objectives, the last is: to evaluate the theoretical properties of the system based on the correctness and its speed

Methodology:
- dito mapupunta yung framework and APIs
- say we will be developing an information system. And all information systems have GET,POST,PUT,DELETE functionalities.

Scope
- 2 things: 
- 1) create a theoretical design for a proposed system
- 2) evaluate the theoretical running time and speed of the operations.
- define here what you mean by proposed system and base system (FTP protocol of NCBI be specific that this is what base system means)
- base system: FTP retrieval from NCBI

Limitation
- no implement
- be transparent, time constraints and pandemic
- no security and authentication

Theoretical Framework
- discusses abstractly anong paglalaruan na ideas
- Discuss Hybrid,CS, P2P in short sentences
- Network, forego layers 1-3. Briefly discuss layers 4 and 5 and their importance.
- Discuss how retrieval operations rely on layers 4 and 5 (FTP for NCBI, torrent??)
- good flow, but too many info, need to shorten it.

Summary of the System
- call this section caDDS
- start with an outline for the entire section
- The retrieval mehanism is inspired by SeqTorr. This section describes the proposed system. This is a hybrid focused or distributed system

Community
- this section is good
- move to 7.2 or after

System Description
- for each photo (GET, POST...) , need one paragraph explaining each one. 1 paragraph per operation (GET, POST...)
- for each photo outline what happens in 1) 2) 3)

Proof of Correctness (BIGTIME CHANGE)
- how do you guarantee that a get does what it is intended to do:
- cite preconditions, post conditions,all the paths from pre conditions to post conditions and how it would always be true
- like an if else
- shows the path from pre condition to post condition
- set pre conditions, assumptions, then show how it becomes postconditions (state after)

Sample:
Assume do not factor in physical state of the network, it is error free
for GET
input of this operation: search key
output of this operation: data transferred to the user

Master node during search has 2 possible output
1) at least 1 hit
2) not in the database

After having at least 1 hit
-> sends the link to user
-> user clicks to link if he wants
-> user presses link
-> user makes a download request to master
-> master redirects user to data node
-> data node sends data to user

Precondition: User chooses to download
Postcondition: User retrieves files

User inputs:search key
master outputs: results report

How to know master will send the correct file?
1) each metadata has a unique identifier
2) guarantees exact file since each file has a unique identifier

How to know it will reach the correct user
1) Each user has a unique identifier (IP Address)

Theoretical Analysis
- Mention base system means FTP from NCBI...
- Mention why only get/ retrieval is used. Since all operations can be broken up into small retrieval operations
- abstract size of a payload
- cite that network is assumed to be error free
- articulate assumptions better
- move the final comparisons here

Conclusion Template:
- say what you did
- summarize analysis of results
- 2 to 3 sentences of what you did/ problem
- 2 sentences correctness and speed
- correctness: system operations
- speed: faster/slower
- shown to do what is expected based on preconditions

Reccomendations
- okay

Acknowledgements -> Remove from Appendix

Biblio
- Make it APA format